\chapter*{Acknowledgements}
I want to express my gratitude to the people who helped me throughout this thesis. Their support and guidance made a huge difference, and I am genuinely thankful for their assistance.

First and foremost, I am extremely grateful for my esteemed supervisors, Christopher Morris and Luis Müller. Together, they formed a great team that not only introduced me to this captivating research field but also provided me with a research question that aligns perfectly with my passion. Their valuable insights, constant encouragement, and belief in my abilities have been the driving force behind this thesis.

In particular, I want to thank Christopher for his dedication and generosity in sharing his time and knowledge despite his commitments as a full-time professor. His responsiveness to my questions, provision of supplementary materials, and flexibility for meetings were invaluable throughout the entire thesis.

Additionally, I want to thank Luis Müller for being a sort of mentor for me. His guidance on best practices for empirical investigations in this research field, creative ideas for experiment setups along with his thorough proofreading and constructive feedback, have significantly enhanced the academic rigor of this work.

The most fruitful moments of this journey were undoubtedly the countless meetings with both Christopher and Luis. Through these discussions, we explored the progress of my thesis, addressed theoretical limitations in the proofs, and delved into intermediate empirical results and their interpretations. Each interaction left me with renewed motivation, making the overall pursuit of this thesis a challenging yet profoundly enjoyable experience.

I also want to thank the RWTH Aachen High Performance Computing cluster for providing the necessary computational resources for my experiments. Without these resources, the research goals of my thesis would not have been possible.

Lastly, I want to thank my parents and family for their unwavering support throughout my studies. Their encouragement means a lot to me and has been a great source of motivation.