Give a short Outline

\section{Setup}
Introduction to this section!

\subsection{Choice of Datasets}
Luis' Taxonomy, verschiedene Bereiche -> TUDataset

\subsection{Choice of Models}
GIN weil sehr expressive \cite{Xu2018}, GAT und GCN wegen \cite{Nikolentzos2023} mit basic pool um complexity der models zu containen und bessere vergleiche zu haben

$\wlnn$ auch nur mit basic pool

\subsection{Experimental Setup}
python 3.10, pytorch und pytorch-geometric (use footnotes from outline). Use of Wheights and Biases as data logging instance, HPC, Colab und meine eigene Hardware - > summierte overall computation time. Standard procedure, training with 10-fold and repition and a split of training, validation and test dataset.

\section{Results}
Introduction to this section!

\subsection{Explain how results look like}
Shortly explain, their performance and how they typically learn. Grafiken über das mean train\_acc, val\_acc, test\_acc vs. epochs.

\subsubsection{1-WL+NN}
Explain most important parameters such as k-wl mit der wl accuracy on. Show how it relates to the datasets by showing the table of theoretical accuracies, 
Grafiken über das mean train\_acc, val\_acc, test\_acc vs. epochs.

\subsubsection{GNN}
Show performance of different GNN models -> Boxplot of different pooling functions
Grafiken über das mean train\_acc, val\_acc, test\_acc vs. epochs.

\subsection{Test Accuracy}
Show a big table with all test-accuracies and its standard deviaton and highlight in bold face, the best one.
Asses the std of both model types.

\subsection{Overfitting Behaviour}
Compare the difference between training data performance vs test accuracy accross both models

\subsection{Aggregate Analysis}
Explain what aggregates are and show tSNE with the of the best models with SVM

Also use KNN to assess how good the actuall aggregates are differentiable. -> local clusters ?

\section{Discusssion}
\subsection{Learned Lessons}
\subsection{Future Work}

\section{Conclusion}